visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports

### 2024-03-08 manuscript.Rmd
revision 2

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_pdf
appendix          :
  - "appendix_1.Rmd"
  - "appendix_2.Rmd"
  - "appendix_3.Rmd"
  - "appendix_4.Rmd"
  - "appendix_5.Rmd"
  - "appendix_6.Rmd"
---

@@

The `bootstrap_rr` function in *visualizemi* was designed to estimate the likely replication of overall model invariance with the assumption that the data used for the estimation represents the larger population. The following arguments are used:

@@

```{r boot-rr, echo = T, eval = F, size = "small"}
boot.model.invariant <- 
  bootstrap_rr(
    # saved configural model 
    saved_configural = results.invariant$model_configural,
    # dataframe
    data = df.invariant,
    # model syntax
    model = model.overall, 
    # group variable column in dataframe
    group = "group",
    # number of bootstraps
    nboot = 1000, 
    # which fit index you would like to use
    invariance_index = "cfi",
    # what is your criterion for that fit index
    invariance_rule = .01,
    # what equality constraints are you testing 
    group.equal = c("loadings", "intercepts", "residuals")
  )
```

@@

$$h_{nmi} = 2 \times (asin\sqrt{p_{data}} - asin\sqrt{p_{random}})$$
The alternative, $h_{mi}$, for effect size of measurement invariance replication would simply be the inverse sign of $h_{nmi}$ and is also included in the table. Two additional columns $h_{nmi_p}$ and $h_{nmi_p}$ represent the $h$ values divided by the upper bound of $h$ (i.e., $\pi$), to help with interpretation of the effect size (thus, bounding $h$ to -1 to 1).

@@

# examine the loadings 
partial.rs.2$model_coef %>% 
  filter(term == "RS6 ~1 ") %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(term, group, estimate, std.error)
# examine the fit indices 
partial.rs.2$model_fit %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(AIC, BIC, cfi, tli, rmsea, srmr)

@@

```{r model-boot-RS, include = T, echo = T, results = "small", eval = F}
boot.model.rs <- 
  bootstrap_rr(
    saved_configural = results.rs$model_configural,
    data = DF,
    model = model.rs, 
    group = "sample", 
    nboot = 1000, 
    invariance_index = "cfi",
    invariance_rule = .01, 
    group.equal = c("loadings", "intercepts", "residuals")
)
```

@@

In summary, if one were planning a replication, the prediction would be that item intercepts would likely not replicate, with a large effect size (i.e., it is easy to judge $h_{nmi_p}$ close to the max of one as large). While this study found partial invariance by relaxing constraints on two individual items, bootstrapped partial invariance indicates that any item could potentially be problematic with an effect size averaging $d$ ~ 0.50 difference in means. While $d_{MACS}$ values represented a "small" effect based on previous publications, this effect may be muted by examining both loadings and intercepts. The results here suggest that the effect is driven by intercepts. The overall average score on items is high: $M_M$ = `r apa_num(mean(apply(DF[ , -ncol(DF)], 2, mean)))` ($M_{SD}$ = `r apa_num(mean(apply(DF[ , -ncol(DF)], 2, sd)))`). Given the mean standard deviation, a $d$ ~ 0.50 represents `r apa_num(mean(apply(DF[ , -ncol(DF)], 2, sd))/2)` or nearly one whole point on the scale. A researcher could decide that at least $d$ = 0.33 or at least a third of a standard deviation would be an important change and set that as their smallest effect size of interest for invariance. Further, a newly planned study should investigate what variables may predict when and why samples separate into bimodal representations for item means.

@@

@@

In this tutorial, we examined how to use multiple tools to examine measurement invariance and its potential replication. Model fit comparisons and statistics can be paired with the proposed effect size measures, and a visualization to examine individual items and the overall latent mean scores. The impact of potential replication was estimated on the overall model and the individual parameters. Using real data, the effect of two non-invariant item intercepts was examined and visualized. This tutorial manuscript has provided a concrete way to plan for pre-registration and/or registered reports. Researchers could simulate results based on published or previously collected data to determine the likelihood and size of potential replication. They could plan and pre-register a smallest effect of interest. For example, we may determine that an $h_{nmi_p}$ value above .20 represents an important level of non-invariance for our model overall, while $h_{nmi_p}$ > .30 for any individual parameter warrant caution against invariance for groups. Others have begun to discuss the importance of focusing on effects in the scale of the data and their practical importance [@cumming2012; @anvari2021].

From the example, our interpretation may be that the difference between group's latent means is large, as a `r apa_num(partial.rs.2$model_coef %>% filter(term == "RS ~1 ") %>% slice_tail() %>% pull(estimate))` change on a 7 point scale is approximately 10% more resiliency for students when compared to the clinical sample. Practically, 10% in resiliency for an area of the United States (Mississippi) often hit with natural disasters (hurricanes, tornadoes, floods) and high levels of poverty would be very important. Even the smaller difference of .5 point on each individual item could translate into increases in resiliency, and these results may elucidate avenues for further exploration into areas of focus within resiliency, given the items.

What do the results of a study on measurement invariance with these results tell us about replication, generalizability, and validity overall? If a researcher decides their effects are large, they should likely caution against suggesting that these scores are directly comparable without weighting or other adjustment. Let's consider a scenario wherein the change metric between models picked (i.e., $\Delta$CFI, $\Delta$RMSEA) indicates a "significant" change in model fit. However, if both the effect size and a visual inspection of the invariance indicates a small difference, we may decide to lessen the practical importance of those results, much like "just significant" *p*-values with small effect sizes are treated now. Given that the goal of measurement invariance is to compare *estimates*, we should expect some differences across samples due to the nature of sampling and estimation. It may be that many of the published models presented represent these effects - small variations between groups due to sampling error or other small crud - but do not represent a fundamental problem with the measurement or generalizability of the results.

@@

# (APPENDIX) Appendix {-}

@@

>>>>

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_pdf
appendix          :
  - "appendix_1.Rmd"
  - "appendix_2.Rmd"
  - "appendix_3.Rmd"
  - "appendix_4.Rmd"
  - "appendix_5.Rmd"
  - "appendix_6.Rmd"
  - "appendix_7.Rmd"
---

@@

The `bootstrap_model` function in *visualizemi* was designed to estimate the likely replication of overall model invariance with the assumption that the data used for the estimation represents the larger population. The following arguments are used:

@@

```{r boot-rr, echo = T, eval = F, size = "small"}
boot.model.invariant <- 
  bootstrap_model(
    # saved configural model 
    saved_configural = results.invariant$model_configural,
    # dataframe
    data = df.invariant,
    # model syntax
    model = model.overall, 
    # group variable column in dataframe
    group = "group",
    # number of bootstraps
    nboot = 1000, 
    # which fit index you would like to use
    invariance_index = "cfi",
    # what is your criterion for that fit index
    invariance_rule = .01,
    # what equality constraints are you testing 
    group.equal = c("loadings", "intercepts", "residuals")
  )
```

@@

$$h_{nmi} = 2 \times (asin\sqrt{p_{data}} - asin\sqrt{p_{random}})$$

The alternative, $h_{mi}$, for effect size of measurement invariance replication would simply be the inverse sign of $h_{nmi}$ and is also included in the table. Two additional columns $h_{nmi_p}$ and $h_{nmi_p}$ represent the $h$ values divided by the upper bound of $h$ (i.e., $\pi$), to help with interpretation of the effect size (thus, bounding $h$ to -1 to 1).

@@

# examine the loadings 
partial.rs.2$model_coef %>% 
  filter(term == "RS6 ~1") %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(term, group, estimate, std.error)
# examine the fit indices 
partial.rs.2$model_fit %>% 
  filter(model == "intercepts") %>% 
  dplyr::select(AIC, BIC, cfi, tli, rmsea, srmr)

@@

```{r model-boot-RS, include = T, echo = T, results = "small", eval = F}
boot.model.rs <- 
  bootstrap_model(
    saved_configural = results.rs$model_configural,
    data = DF,
    model = model.rs, 
    group = "sample", 
    nboot = 1000, 
    invariance_index = "cfi",
    invariance_rule = .01, 
    group.equal = c("loadings", "intercepts", "residuals")
)
```

@@

In summary, if one were planning a replication, the prediction would be that item intercepts would likely not replicate, with a large effect size (i.e., it is easy to judge $h_{nmi_p}$ close to the max of one as large). While this study found partial invariance by relaxing constraints on two individual items, bootstrapped partial invariance indicates that any item could potentially be problematic with an effect size averaging $d$ \~ 0.50 difference in means. While $d_{MACS}$ values represented a "small" effect based on previous publications, this effect may be muted by examining both loadings and intercepts. The results here suggest that the effect is driven by intercepts. The overall average score on items is high: $M_M$ = `r apa_num(mean(apply(DF[ , -ncol(DF)], 2, mean)))` ($M_{SD}$ = `r apa_num(mean(apply(DF[ , -ncol(DF)], 2, sd)))`). Given the mean standard deviation, a $d$ \~ 0.50 represents `r apa_num(mean(apply(DF[ , -ncol(DF)], 2, sd))/2)` or nearly one whole point on the scale. A researcher could decide that at least $d$ = 0.33 or at least a third of a standard deviation would be an important change and set that as their smallest effect size of interest for invariance. Further, a newly planned study should investigate what variables may predict when and why samples separate into bimodal representations for item means.

@@

## An Example Extension

One benefit of the open science movement on scale development is the publication of datasets or covariance tables with published articles. We can extend our examination of potential replication on other variables that may effect assessment of underlying phenomena. For example, scale translation across languages is not only impacted by the literal conversion of concepts from one language to another, but also the cultural and societal norms of the target population [@cha2007; @yu2004; @chang1999]. The RS14 was tested in Chinese speaking samples in @chen2020 across five different large samples and determined that the scale showed good psychometric properties for use within Chinese speaking samples as a one-factor model of resiliency. Given these results, another researcher may assume that the models would be comparable between English speaking (i.e., likely United States) and Chinese speaking samples. With the published data, we can use `visualizemi` to determine if the RS14 is invariant across language/culture, and if the results would likely replicate if tested on a new set of samples, and what, if any, effect size differences are found in parameters. The code used for this analysis is presented in Appendix 7.

The @aiena2014 data used in our previous example was first filtered to only college students, as we have already noted that clinical samples show slightly different intercepts for at least two of the items from student samples. The @chen2020 data also included college students, which allows us to test a comparable sample that varies on translation and culture. The test of measurement invariance indicated that the factor structure and loadings were invariant across groups. Much like our test of clinical versus student samples, the results indicated that the intercepts were not consistent across groups. Within the English clinical/student sample, partial invariance could be achieved by relaxing two item intercept constraints (item 6 and 7). To achieve "partial" invariance between the Chinese and English speaking samples, we would need to relax more than half of the items (specifically, eight items: 1, 2, 3, 4, 10, 11, 12, 14), and it would be difficult to suggest partial invariance given this finding. $d_{MACS}$ values range from 0.26 to 0.32 for the eight items which could be interpreted a small to medium given @nye2019's simulation study. Figure \@ref(fig:rs14-chinese-pic) portrays the results from the second item on the RS14 (*life accomplishments*).

The results of model bootstrapping indicated that the effect of the loadings was likely to replicate (only invariant results were found), but the intercepts were never found to be invariant compared to a randomized sample ($h_{nmi_p}$ = 1). Therefore, we would expect that these differences are persistent, either due to the adaptation or cultural differences in resiliency across samples. The bootstrapped partial invariance demonstrates that each item intercept has a medium to large difference between the two samples as shown in Figure \@ref(fig:rs14-chinese-effect-pic), which may explain why full or partial invariance cannot be achieved. This result does not invalidate either version of the scale, but informs researchers of potential baseline differences in item responding for the two samples. Therefore, careful interpretation should be made when comparing these two samples in other instances, as differences in latent means may be the default finding, but with these results one may determine if their results are different from expectations of general scale responding.

```{r rs14-chinese-pic, include = T, fig.cap = "The differences in intercepts for the second item on the Rs14 by language sample. The differences between intercepts are shown on the left hand side with a clear separation between lines. The latent means also show a clear difference between groups where the English group appears to have lower scores overall than the Chinese group."}
knitr::include_graphics("figures/rs14-chinese-RS2.png")
```

```{r rs14-chinese-effect-pic, include = T, fig.cap = ""}
knitr::include_graphics("figures/rs14-chinese-effect_invariance_plot.png")
```

@@

In this tutorial, we examined how to use multiple tools to examine measurement invariance and its potential replication. Model fit comparisons and statistics can be paired with the proposed effect size measures, and a visualization to examine individual items and the overall latent mean scores. The impact of potential replication was estimated on the overall model and the individual parameters. Using real data, the effect of two non-invariant item intercepts was examined and visualized. This tutorial manuscript has provided a concrete way to plan for pre-registration and/or registered reports. Researchers could simulate results based on published or previously collected data to determine the likelihood and size of potential replication. They could plan and pre-register a smallest effect of interest. For example, we may determine that an $h_{nmi_p}$ value above .20 represents an important level of non-invariance for our model overall, while $h_{nmi_p}$ \> .30 for any individual parameter warrant caution against invariance for groups. Others have begun to discuss the importance of focusing on effects in the scale of the data and their practical importance [@cumming2012; @anvari2021].

From the example, our interpretation may be that the difference between group's latent means is large, as a `r apa_num(partial.rs.2$model_coef %>% filter(term == "RS ~1 ") %>% slice_tail() %>% pull(estimate))` change on a 7 point scale is approximately 10% more resiliency for students when compared to the clinical sample. Practically, 10% in resiliency for an area of the United States (Mississippi) often hit with natural disasters (hurricanes, tornadoes, floods) and high levels of poverty would be very important. Even the smaller difference of .5 point on each individual item could translate into increases in resiliency, and these results may elucidate avenues for further exploration into areas of focus within resiliency, given the items. The secondary example showed that we can extend these results to other samples to examine other potentially impactful variables on assessment. The findings replicate in the sense that the scale shows the same invariant issues with intercepts on a Chinese versus English sample comparison. However, in this analysis, it is clear that the differences in items averages are much larger and across all items, rather than a few.

What do the results of a study on measurement invariance with these results tell us about replication, generalizability, and validity overall? If a researcher decides their effects are large, they should likely caution against suggesting that these scores are directly comparable without weighting or other adjustment. Let's consider a scenario wherein the change metric between models picked (i.e., $\Delta$CFI, $\Delta$RMSEA) indicates a "significant" change in model fit. However, if both the effect size and a visual inspection of the invariance indicates a small difference, we may decide to lessen the practical importance of those results, much like "just significant" *p*-values with small effect sizes are treated now. The results from our Chinese versus English comparison show us the other scenario: non-invariant results that clearly indicate differences with a large effect size on both replication and item average differences. Given that the goal of measurement invariance is to compare *estimates*, we should expect some differences across samples due to the nature of sampling and estimation. It may be that many of the published models presented represent these effects - small variations between groups due to sampling error or other small crud - but do not represent a fundamental problem with the measurement or generalizability of the results. The `visualizemi` package is one useful tool to help sort out if findings are this small sampling error or differences in samples due to other variables.

@@

# (APPENDIX) Appendix {.unnumbered}

@@

```{r child = "appendix_7.Rmd"}
